Finch: Sparse and Structured Array Programming with Control Flow

### 2024-03-26 sections/eval.tex
Update on Overleaf.

Sparse Matrix-Matrix Multiplication (SpGEMM) is a fundamental operation in scientific computing and data analytics. SpGEMM focuses on efficiently computing the product of two sparse matrices to produce an output sparse matrix.
The inner product, outer product, and Gustavson's algorithm are three different approaches to performing SpGEMM. 
Inner product involves multiplying two matrices by taking dot products of corresponding rows and columns.
Outer product computes the product of all pairs of elements from the input matrices.

More detailed explanation is available in Section 2.2 of ~\cite{} \changwan{gamma}.

@@

\help{6. explain the features of "datastructure-driven programming" that are used here in this problem. i.e. explicit temporary, bytemap format, transpositions}.
\begin{figure}
	\includegraphics[width=\linewidth]{spgemm_small_speedup_log_scale.png}
    \caption{A comparison of several matrix multiplication algorithms between Finch and Taco smaller matrices, ordered from small to big dimension. Note that inner products necessarily requires $O(n^2)$ work and taco's outer products format is dense. Finch can use a sparse outer products format and thus has an asymptotic advantage that becomes evident as the output dimensions grow.}
\end{figure}

\begin{figure}
	\includegraphics[width=\linewidth]{spgemm_joel_speedup.png}
    \caption{A comparison of gustavson's algorithm between Finch and Taco on some larger matrices}
\end{figure}

@@

\subsubsection{Finch Interpreter} 
Once the Finch Logic program has been converted to the standard form, the Finch Interpreter can execute each query, in order, through a straightforward lowering process. First, the output format is identified by unpacking the outer $\finchreformat$ statement. Next, the inner expression of the is unpacked to identify the aggregate operator and to convert the pointwise expression into a Finch expression. At this point, any aliases to the result of previous queries are replaced with an access to the actual tensor. Lastly, the concordant loop order is identified and instantiated. At this point, the lowered query can be compiled and executed with the Finch compiler, and the result is assigned to $name$ in the plan's scope before proceeding to the next query in the plan.

>>>>

Sparse Matrix-Matrix Multiplication (SpGEMM) is a fundamental operation in scientific computing and data analytics. SpGEMM focuses on efficiently computing the product of two sparse matrices to produce an output sparse matrix.
The inner product, outer product, and Gustavson's algorithm are three different approaches to performing SpGEMM. 
Inner product involves multiplying two matrices by taking dot products of corresponding rows and columns.
Outer product computes the product of all pairs of elements from the input matrices.
More detailed explanation is available in Section 2.2 of ~\cite{} \changwan{gamma}.
\changwan{Inner, asym bad, }

Figure~\ref{spgemm_small} 
TACO does not support outer product due to its lack of support for scatter into a 2D workspace.

@@

\help{6. explain the features of "datastructure-driven programming" that are used here in this problem. i.e. explicit temporary, bytemap format, transpositions}.
\begin{figure}
    \label{spgemm_small}
	\includegraphics[width=\linewidth]{spgemm_small_speedup_log_scale.png}
    \caption{A comparison of several matrix multiplication algorithms between Finch and Taco smaller matrices, ordered from small to big dimension. Note that inner products necessarily requires $O(n^2)$ work and taco's outer products format is dense. Finch can use a sparse outer products format and thus has an asymptotic advantage that becomes evident as the output dimensions grow.}
\end{figure}

\begin{figure}
\label{spgemm_large}
	\includegraphics[width=\linewidth]{spgemm_joel_speedup.png}
    \caption{A comparison of gustavson's algorithm between Finch and Taco on some larger matrices}
\end{figure}

@@

\subsubsection{Finch Interpreter} 
Once the Finch Logic program has been converted to the standard form, the Finch Interpreter can execute each query, in order, through a straightforward lowering process. First, the output format is identified by unpacking the outer $\finchreformat$ statement. Next, the inner expression of the is unpacked to identify the aggregate operator and to convert the pointwise expression into a Finch expression. At this point, any aliases to the result of previous queries are replaced with an access to the actual tensor. Lastly, the concordant loop order is identified and instantiated. The lowered query can then be compiled and executed with the Finch compiler, and the result is assigned to $name$ in the plan's scope before proceeding to the next query in the plan.
